---
title: "How does carrying capacity vary with temperature?"
output: 
  html_document: 
    fig_caption: yes
    keep_md: yes
    toc: yes
    theme: "lumen"
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(broom)
library(gridExtra)
library(lubridate)
library(plotrix)
library(stringr)
library(nlstools)
library(simecol)
library(purrr)
```

### Background

#### We asked

_How does carrying capacity vary with temperature?_

####  We predicted

* Carrying capacity should decrease with warming, due to higher metabolic demands. As such, the temperature depence of carrying capacity should vary inversely with the temperature dependence of photosynthesis (e.g. Ea = 0.32 eV), which limits metabolism in autotrophs. 

#### Our goals for now:

* estimate r and K by fitting a logistic growth model to our experimental data

* estimate the temperature dependence of K (i.e. Ea of K)

* estimate the activation energies of photosynthesis and respiration in T. Tetrahelis


### Estimate activation energies for photosynthesis and respiration

Since we are testing the prediction that carrying capacity scales with the inverse of the activation energy of photosynthesis in autotrophs (Savage et al. 2004), we first need to estimate the Ea of photosynthesis. We do that here empirically with T. Tetrahelis by measuring rates of oxygen flux in the light and dark in the lab. 

*Note! these are practice respirometry data from Scenesdesmus obliquus...I expect the T. Tetrahelis data to look a little different*

```{r message = FALSE, warning= FALSE}
metabolic_rate_data <- read_csv("/Users/Joey/Documents/J-TEMP/data-raw/scenedesmus_photo_resp_rates.csv")

metabolic_rate_data %>%
	gather(key = flux_type, value = rate_estimate, GPP, contains("corrected")) %>% 
	mutate(rate_estimate = rate_estimate * 3600) %>% 
	group_by(temperature.x, flux_type) %>% 
	summarise_each(funs(mean, std.error), rate_estimate) %>%
	mutate(mean = ifelse(grepl("respiration", flux_type), mean*-1, mean)) %>%
	mutate(flux_type = ifelse(grepl("corrected_respiration_slope", flux_type), "respiration", flux_type)) %>% 
	mutate(flux_type = ifelse(grepl("corrected_photosynthesis_slope", flux_type), "photosynthesis", flux_type)) %>%
	# mutate(mean = ifelse(grepl("respiration", flux_type), mean*-1, mean)) %>% 
	ggplot(aes(x = temperature.x, y = mean)) + geom_point(size = 3) +
	facet_wrap( ~ flux_type) + geom_errorbar(aes(ymin = mean - std.error, ymax = mean + std.error), width = 0.1) +
	xlab("temperature") +
	ylab("oxygen flux (mg/L*hour)") + geom_hline(yintercept = 0) + theme_bw() 
```

Plots to get activation energies
```{r, echo=FALSE}
metabolic_rate_data %>% 
	gather(key = flux_type, value = rate_estimate, GPP, contains("corrected")) %>% 
	mutate(rate_estimate = rate_estimate * 3600) %>% 
	mutate(temperature.x = as.numeric(temperature.x)) %>% 
	mutate(inverse_temp = (1/(.00008617*(temperature.x+273.15)))) %>%
	mutate(rate_estimate = ifelse(grepl("respiration", flux_type), rate_estimate*-1, rate_estimate)) %>%
	mutate(flux_type = ifelse(grepl("corrected_respiration_slope", flux_type), "respiration", flux_type)) %>% 
	mutate(flux_type = ifelse(grepl("corrected_photosynthesis_slope", flux_type), "photosynthesis", flux_type)) %>% 
	# filter(flux_type == "GPP") %>% 
	ggplot(aes(x = inverse_temp, y = log(rate_estimate))) + geom_point(size = 4, color = "blue", alpha = 0.5) +
	facet_wrap( ~ flux_type) +
	geom_smooth(method = lm) +
	scale_x_reverse() + theme_bw() + ylab("ln metabolic rate (mg O2/L*hr") + xlab("inverse temperature (1/kT)") 
```

Estimate activation energies
```{r echo = FALSE}
metabolic_rate_data %>% 
	gather(key = flux_type, value = rate_estimate, GPP, contains("corrected")) %>% 
	mutate(rate_estimate = rate_estimate * 3600) %>% 
	# group_by(temperature.x, flux_type) %>% 
	# summarise_each(funs(mean, std.error), rate_estimate) %>% 
	# ungroup() %>% 
	mutate(temperature.x = as.numeric(temperature.x)) %>% 
	mutate(inverse_temp = (1/(.00008617*(temperature.x+273.15)))) %>% 
	mutate(rate_estimate = ifelse(grepl("respiration", flux_type), rate_estimate*-1, rate_estimate)) %>%
	mutate(flux_type = ifelse(grepl("corrected_respiration_slope", flux_type), "respiration", flux_type)) %>% 
	mutate(flux_type = ifelse(grepl("corrected_photosynthesis_slope", flux_type), "photosynthesis", flux_type)) %>% 
	group_by(flux_type) %>% 
	do(tidy(lm(log(rate_estimate) ~ inverse_temp, data = .), conf.int = TRUE)) %>%
	filter(term != "(Intercept)") %>% 
	knitr::kable(., align = 'c', format = 'markdown', digits = 2)

```




Read in time series population dynamics data
```{r message=FALSE, warning=FALSE}
sea <- read_csv("/Users/Joey/Documents/J-TEMP/data-processed/sea_processed.csv")
	all_cells_TT <-	sea %>%
		filter(species == "TT") %>% 
		unite(Unique_ID, temperature, rep, remove = FALSE) %>%  
		rename(replicate = rep) %>% 
		rename(P = cell_density,
					 days = time_since_innoc_days,
					 ID = Unique_ID) %>% 
		arrange(ID, days) %>% 
		filter(!grepl("NA", ID)) 
```

### Logistic growth model

Define our logisitic growth model
```{r}

Parameters <- c(r = 1, K = 10 ^ 7) ## initial parameter guesses
CRmodel <- new("odeModel",
							 main = function (time, init, parms) {
							 	with(as.list(c(init, parms)), {
							 		dp <-  r * P * (1 - (P / K))
							 		list(c(dp))
							 	})
							 },
							 parms = Parameters,
							 times = c(from = 0, to = 43, by = 0.1), # the time interval over which the model will be simulated.
							  init = c(P = 2000),
							 # init = c(P = 971410.2), # starting biovolume
							 solver = "lsoda" #lsoda will be called with tolerances of 1e-9. Default tolerances are both 1e-6. Lower is more accurate.
)

fittedparms <- c("r", "K") # for assigning fitted parameter values to fittedCRmodel

controlfit <- function(data){
	
	init(CRmodel) <- c(P = data$P[1]) # Set initial model conditions to the biovolume taken from the first measurement day
	obstime <- data$days # The X values of the observed data points we are fitting our model to
	yobs <- select(data, P) # The Y values of the observed data points we are fitting our model to
	
	
	fittedCRmodel <- fitOdeModel(CRmodel, whichpar = fittedparms, obstime, yobs,
															 debuglevel = 0, fn = ssqOdeModel,
															 method = "PORT", lower = LowerBound, upper = UpperBound, scale.par = ParamScaling,
															 control = list(trace = T)
	)
	
	r <- coef(fittedCRmodel)[1]
	K <- coef(fittedCRmodel)[2]
	ID <- data$ID[1]
	output <- data.frame(ID, r, K)
	return(output)
}

```

Define the function that we will use for generating our model fit plots

```{r}
plotsinglefit <- function(data){
	
	init(CRmodel) <- c(P = data$P[1]) # Set initial model conditions to the biovolume taken from the first measurement day
	obstime <- data$days # The X values of the observed data points we are fitting our model to
	yobs <- select(data, P) # The Y values of the observed data points we are fitting our model to
	# parms(CRmodel)[TempName] <- data$temp[1] # Set the temperature parameter in CRmodel to whatever our control replicate used.
	
	# Below we fit a CRmodel to the replicate's data. The optimization criterion used here is the minimization of the sum of
	# squared differences between the experimental data and our modelled data. 
	
	# The PORT algorithm is used for the model fitting, analogous to O'Connor et al.
	# "lower" is a vector containing the lower bound constraints
	# for the parameter values.
	
	fittedCRmodel <- fitOdeModel(CRmodel, whichpar = fittedparms, obstime, yobs,
															 debuglevel = 0, fn = ssqOdeModel,
															 method = "PORT", lower = LowerBound, upper = UpperBound, scale.par = ParamScaling,
															 control = list(trace = T)
	)
	
	# To display the fitted results we need to create a new OdeModel object. Here
	# we duplicate CRmodel and then alter it to use our new fitted parameters.
	plotfittedCRmodel <- CRmodel
	parms(plotfittedCRmodel)[fittedparms] <- coef(fittedCRmodel)
	
	# set model parameters to fitted values and simulate again
	times(plotfittedCRmodel) <- c(from=0, to=43, by=1)
	ysim <- out(sim(plotfittedCRmodel, rtol = 1e-9, atol = 1e-9))
	
	# Form observed data into a dataframe; the simulated data are already in a dataframe
	observeddata <- data.frame(obstime, yobs)
	observeddata$type <- "observed"
	observeddata <- rename(observeddata, time = obstime)
	simulateddata <- ysim
	simulateddata$type <- "simulated"
output <- bind_rows(observeddata, simulateddata) ## plop everything together into one data frame
	

	# Alternative: Plot the results of our model fitting directly.
	# biol_plot <- ggplot() +
	# 	geom_point(data = observeddata, aes(x = obstime, y = yobs, color = "observed")) + # Observed data are points
	# 	geom_line(data = simulateddata, aes(x = time, y = P, color = "simulated")) + # Simulated data are in a continuous line
	# 	labs(x = "Time (days)", y = "Algal Biovolume")
	
	# Output the results as a ggplot2 object
	# output <- biol_plot
	return(output)
}
```

### Fitting process

Set upper and lower bounds and initial guesses for our parameters
```{r}
Parameters <- c(r = 1, K = 10 ^ 7)
Parameters <- c(r = 1, K = 10 ^ 5)
# Declare the parameters to be used as the bounds for the fitting algorithm
LowerBound <- c(r = 0.01, K = 10 ^ 2)
UpperBound <- c(r = 10, K = 10 ^ 7) 

# Declare the "step size" for the PORT algorithm. 1 / UpperBound is recommended
# by the simecol documentation.
ParamScaling <- 1 / UpperBound
```

Onto the actual fitting! The output_TT will be a dataframe with all of our estimated r and Ks.
```{r, eval=FALSE}
#Going to cheat here and not run this here. Since it'll take a while! But I'm including it so we can see it.

TT_split <- split(all_cells_TT, f = all_cells_TT$ID) ## split the TT dataframe into little mini dataframes, one for each replicate - temperature combination. We will then use a map function to apply our fitting function to each dataframe individually.

output_TT <- TT_split %>% 
	map_df(controlfit)

write_csv(output_TT, "/Users/Joey/Documents/J-TEMP/data-processed/output_rK_TT.csv")
write_csv(output_TT, "/Users/Joey/Documents/J-TEMP/data-processed/output_rK_TT_cell_abundance.csv")

## just for now, rename the output_TT as output_TT_abundance
output_TT_abundance <- output_TT
```


### Plot the model fits
```{r, eval=FALSE}
obs_sim_data <- TT_split %>% 
	map_df(plotsinglefit, .id = "ID") ## don't run this here

obs_sim_data %>% 
	separate(ID, into = c("temperature", "rep"), remove = FALSE) %>% 
	mutate(temperature = as.numeric(temperature)) %>% 
	mutate(type = str_replace(type, "simulated", "fit"))

# write_csv(obs_sim_data, "/Users/Joey/Documents/J-TEMP/data-processed/TT_obs_sim_data.csv")
write_csv(obs_sim_data, "/Users/Joey/Documents/J-TEMP/data-processed/TT_obs_sim_data_cell_abundance.csv")
```


Plot the observed and simulated data, based on our fitted parameters to see how they match up
```{r message=FALSE, warning=FALSE}
obs_sim_data <- read_csv("/Users/Joey/Documents/J-TEMP/data-processed/TT_obs_sim_data.csv")

obs_sim_data %>% 
	filter(ID == "16_5") %>% 
	ggplot(aes(x = time, y = P, color = type)) + geom_point(size = 3)
	


ggplot() +
	geom_point(data = obs_sim_data, aes(x = time, y = P, color = type)) + 
	facet_wrap( ~ ID) + theme_bw() + ylab("biovolume (um3/ml)") + xlab("days")
# ggsave("figures/time_series_fits_TT.png", width = 12, height = 10)

```


### Explore the K - temp relationships
```{r message=FALSE, warning=FALSE}
## cheat here and pull in the already run model outputs
output_TT <- read_csv("/Users/Joey/Documents/J-TEMP/data-processed/output_rK_TT.csv")
```

Plot K vs. temperature (for the temperatures below Topt...32 and 38 are above Topt so omit here)
```{r}
output_TT %>% 
	separate(ID, into = c("temperature", "rep")) %>% 
	mutate(temperature = as.numeric(temperature)) %>% 
	filter(log(K) < 18) %>% 
	# filter(r < 1) %>% ## these are weirdo fits. Will come back to seeing if I can fix this
	# filter(temperature < 38) %>% 
	ggplot(aes(x = temperature, y = log(K))) + geom_point(size = 4, alpha = 0.5, color = "black") + geom_smooth(method = lm, color = "black") + theme_bw() + ylab("ln K (biovolume um3/mL)") + xlab("temperature (C)") + theme(text = element_text(size=20))
ggsave("/Users/Joey/Documents/J-TEMP/figures/k-temp-TT.png", width = 7, height = 5)
ggsave("/Users/Joey/Documents/J-TEMP/figures/k-temp-TT-incl32.png", width = 7, height = 5)
ggsave("/Users/Joey/Documents/J-TEMP/figures/k-cell-abundance-temp-TT-32.png", width = 7, height = 5)
```


Get the activation energy of K (for the temperatures below Topt...32 and 38 are above Topt so omit here)
```{r}
output_TT %>% 
	separate(ID, into = c("temperature", "rep")) %>% 
	mutate(temperature = as.numeric(temperature)) %>% 
	filter(log(K) < 18) %>% 
	# filter(temperature < 35) %>% 
	# filter(K < 10^7) %>%
	# filter(r < 1) %>%
	mutate(inverse_temp = (1/(.00008617*(temperature+273.15)))) %>%
	do(tidy(lm(log(K) ~ inverse_temp, data = .), conf.int = TRUE)) %>% 
	knitr::kable(., align = 'c', format = 'markdown', digits = 2)
```


### At carrying capacity: cell traits

Let's look at what's happening at the final timepoint

```{r message = FALSE, warning = FALSE}
TT_cell_data <- read_csv("/Users/Joey/Documents/J-TEMP/data-processed/CH_TT_chla_biovolume_final_time.csv")

TT_cell_data %>% 
	filter(species == "TT") %>% 
ggplot(aes(x = temperature, y = obs)) + geom_point(size = 4, color = "blue", alpha = 0.5) +
	facet_wrap( ~ type, scales = "free") + theme_bw()
```


### Thermal performance curves

```{r warning = FALSE, message = FALSE}

TT <- read_csv("/Users/Joey/Documents/J-TEMP/data-raw/TT_2015.csv")
TT %>% 
	group_by(Temperature, N.Treatment) %>% 
	do(tidy(nls(Particles.per.ml ~ 75 * (1+a)^(Hours.since.Innoc),
							data= .,  start=list(a=0.01),
							control = nls.control(maxiter=100, minFactor=1/204800000)))) %>%
	ggplot(aes(x = Temperature, y = estimate, color = factor(N.Treatment))) + geom_point(size = 4) +
	geom_line(size = 2) + theme_bw() + facet_wrap( ~ N.Treatment) + ylab("population growth rate")

```

Activation energies for population growth rate, r

Fit using the schoolfield model

```{r warning = FALSE, message = FALSE}

coefs_TT <- read_csv("/Users/Joey/Documents/J-TEMP/data-processed/fitted_r_TT_2015.csv")

ggplot(data = coefs_TT, aes(x = ntreatment, y = estimate)) + geom_point(size = 3) +
	facet_wrap( ~ term, scales = "free") + theme_bw()
```